---
title: "Meta analyses in python"
author: "Ashfaq Ali"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R environment
Setting up an R  environment can be done by coreectly setting up the PATH. If you set up the pyhton using the PATH, it will always be used first and the other commands to set python environment will be ignored. 
 
```{r setup, include=FALSE}
Sys.setenv(RETICULATE_PYTHON = "/Users/cob-aaf/opt/anaconda3/envs/omicsInt/bin/python3")
library(reticulate)
use_condaenv(condaenv = "omicsInt")
```

```{r}
usethis::edit_r_profile() 
usethis::edit_r_environ()
```

```{r}
suppressMessages(suppressWarnings(library(GEOquery)))
suppressMessages(suppressWarnings(library(tidyverse)))

```

```{r}
reticulate::conda_install("numpy")
```


```{python librarries}
import numpy as np
import pandas as pd
```


```{python import}
Liver = pd.read_csv("Data/GSE130970_all_sample_salmon_tximport_counts_entrez_gene_ID.csv", index_col=0)
Liver=pd.DataFrame(Liver)
Liver_copy = Liver.copy()
Liver_copy
Liver_copy.drop(columns= '440349.1.X_1', inplace=True)
np.log2(Liver_copy) 
import seaborn as sbn
sbn.histplot(Liver_copy)
```

```{bash}
cat "Data/GSE130970_all_sample_salmon_tximport_counts_entrez_gene_ID.csv" | head
```

```{python}
from sklearn.decomposition import PCA, IncrementalPCA
import plotly.express as px
import matplotlib.pyplot as plt
n_components=2
ipca = IncrementalPCA(n_components=n_components, batch_size=10)
X_ipca = ipca.fit_transform(Liver_copy)

pca = PCA(n_components=n_components)
X_pca = pca.fit_transform(Liver_copy)

fig = px.scatter(X_pca, x=0, y=1, color=Liver_ )
fig.show()
```


```{python}
colors = ['navy', 'turquoise']
plt.scatter(X_pca[:, 0], X_pca[:,1 ],color=colors, lw=2)
pca_liver = PCA(n_components=2)
principalComponents_breast = pca_breast.fit_transform(x)
```
```{python}

from sklearn.decomposition import PCA

df = px.data.iris()
X = df[['sepal_length', 'sepal_width', 'petal_length', 'petal_width']]

pca = PCA(n_components=2)
components = pca.fit_transform(X)

fig = px.scatter(components, x=0, y=1, color=df['species'])
fig.show()
```

```{python importgeo, warnings=FALSE}
import GEOparse as gp
# gse = gp.get_GEO(filepath="Data/GSE130970_series_matrix.txt")
# gse.database
gse = gp.get_GEO(filepath="Data/GSE130970_family.soft")
Liver_phenopd = pd.DataFrame(gse.phenotype_data)
Liver_phenopd
```


## Regular expressions to manipulate phenotype data

 - Find if a pattern exist
 - Get all instances of a pattern
 - clean the data by string splitting
 
### steps
    - define patterns for matching
    - apply the patterns to strings
    - use results of those patterns in data processing

Functions in re include *match()*, *search()*, *findall()*, *split()* ....

```{python}
import re
# create some text
text = "this is a good day and I am having fun with the python language!, the day is going fine but on another day I could have been happier"

# explore the text

if re.search("bad", text):
    print("awesome")
else:
    print("not")
```

```{python}
# segment the string
re.split("day", text)
```
```{python}
re.findall("day", text)
```

## Patterns

Let's start with anchors. Anchors specify the start and/or end of the string that you're trying to match.
The caret (^) character means Start, and the dollar sign ($) character means End. 

If you put caret before a string, it means that the texts that the regex processor retrieves must start with the string you specify. For ending, you have to put the dollar sign character after the string.

```{python}

if re.search("^this", text):
    print("match found")
else:
    print("match not found")

if re.search("happier$", text):
    print("match found")
else:
    print("match not found")
    



```

## work with characters

```{python}
# lets work with a list of league matches for the coming weekends
league ="ABEBACDABECCCAAA"
re.findall('B', league)
re.findall("[AB]", league)

```

```{python}
". ^ $ * + ? { } 
# Set operator [ ] can use explicit mentions or indices

\ | ( )"


```


```{r}
gds858 <- getGEO(filename='Data/GSE130970_series_matrix.txt')
liverPh <- phenoData(gds858)@data
liverPh
```



## Including Plots
```{python}
Liver.shape
Liver[3]
```
```{python}
from PIL import Image
from IPython.display import display
```
```{python image}
hus_img = Image.open("images/house_pic.jpg")
display(hus_img)
```

```{python}
hus_array = np.array(hus_img)
hus_array.shape
hus_array
```

